# LD matrix

```{r}
#| label: setup
#| warning: false
#| message: false
#| echo: false

library(data.table)
library(dplyr)
```

## Data

```{r}
#| label: data

# pQTLs
pqtls <- fread(
    "./data/00_data_pqtls.tsv",
    header = TRUE, data.table = FALSE, sep = "\t"
)

# LD matrix
load("./data/01_data_ldmat.Rda")
```

## LD matrix {#sec-ldmat}

::: {.panel-tabset}

### Sclerostin pQTLs

```{r}
#| label: fig-ld-mat-pqtls
#| fig-cap: LD matrix sclerostin pQTLs
#| out-width: 80%

# LD matrix for pQTLs
ld_mat_pqtls <- ld_mat[
    match(pqtls$rsid, rownames(ld_mat)),
    match(pqtls$rsid, colnames(ld_mat)),
    drop = FALSE
]

# Compute r2 from r
ld_mat_pqtls <- round(ld_mat_pqtls^2, 3)

# Heatmap
heatmaply::heatmaply(
    ld_mat_pqtls,
    label_names = c("Variant 1", "Variant 2", "r2"),
    row_dend_left = TRUE,
    key.title = "r2"
)
```

### Sclerostin pQTLs and *CD300LG* variants

Since *CD300LG* located nearby to *SOST* is known to be associated with HDL cholesterol and triglycerides, we examined whether any of the sclerostin pQTLs are in LD with either rs72836561 (a functional variant in *CD300LG* known to be associated with HDL cholesterol and triglycerides [@surakka2015]) and / or rs72836567 (the top associated variant with *CD300LG* expression in GTEx [@gtex2020]). Two of the sclerostin pQTLs (rs1107747 & rs4793023) are mildly correlated ($r^2$ > 0.05) with rs72836567. So, for these sclerostin pQTLs we used rs72836567 to adjust out any residual association with HDL cholesterol and triglycerides due to the effects of the *CD300LG* gene [@yang2012].  

```{r}
#| label: fig-ld-mat-pqtls-cd300lg
#| fig-cap: LD matrix sclerostin pQTLs and *CD300LG* variants
#| out-width: 60%

# LD matrix for pQTLs and CD300LG variants
ld_mat_pqtls_cd300lg <- ld_mat[
    match(c(pqtls$rsid, "rs72836561", "rs72836567"), rownames(ld_mat)),
    match(c(pqtls$rsid, "rs72836561", "rs72836567"), colnames(ld_mat)),
    drop = FALSE
]

# Compute r2 from r
ld_mat_pqtls_cd300lg <- round(ld_mat_pqtls_cd300lg^2, 3)

# CD300LG correlations
ld_mat_pqtls_cd300lg <- ld_mat_pqtls_cd300lg[
    pqtls$rsid,
    c("rs72836561", "rs72836567"),
    drop = FALSE
]

# Heatmap
heatmaply::heatmaply(
    ld_mat_pqtls_cd300lg,
    dendrogram = "none",
    label_names = c("Sclerostin pQTL", "CD300LG variant", "r2"),
    limits = c(0, 1),
    key.title = "r2"
)
```

:::
