# *SOST* region associations

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: false

# Libraries
library(dplyr)
library(geni.plots)
library(reactable)
library(ggiraph)

# Functions
source("./scripts/00_functions.R")
```

Genetic associations in the *SOST* region for the datasets in [@tbl-gwas-study-info]. [^1]  

```{r}
#| label: data

# Data
load("./data/01_data_ldmat.Rda")
load("./data/02_data_gwas_sost_region.Rda")

# SOST region
gwas <- gwas %>%
  filter(
    chr == "17" &
      pos >= 41831099 - 100000 &
      pos <= 41836156 + 100000
  )
```

## Original datasets

Genetic associations in the *SOST* region for the datasets in [@tbl-gwas-study-info] that were analysed by @zheng2023 (or equivalent[^2]).  

::: {.panel-tabset}

### MI

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-mi
#| fig-cap: Regional plot of the *SOST* region for myocardial infarction (GCST011365)

gwas %>%
  filter(id == "GCST011365") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-mi
#| fig-cap: QQ plot of the *SOST* region for myocardial infarction (GCST011365)

gwas %>%
  filter(id == "GCST011365") %>%
  qq_plot()
```

:::

### CAD

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-cad
#| fig-cap: Regional plot of the *SOST* region for coronary artery disease (GCST005194)

gwas %>%
  filter(id == "GCST005194") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-cad
#| fig-cap: QQ plot of the *SOST* region for coronary artery disease (GCST005194)

gwas %>%
  filter(id == "GCST005194") %>%
  qq_plot()
```

:::

### IS

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-is
#| fig-cap: Regional plot of the *SOST* region for ischemic stroke (GCST005843)

gwas %>%
  filter(id == "GCST005843") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-is
#| fig-cap: QQ plot of the *SOST* region for ischemic stroke (GCST005843)

gwas %>%
  filter(id == "GCST005843") %>%
  qq_plot()
```

:::

### CES

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-ces
#| fig-cap: Regional plot of the *SOST* region for cardioembolic stroke (GCST005842)

gwas %>%
  filter(id == "GCST005842") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-ces
#| fig-cap: QQ plot of the *SOST* region for cardioembolic stroke (GCST005842)
gwas %>%
  filter(id == "GCST005842") %>%
  qq_plot()
```

:::

### T2D

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-t2d
#| fig-cap: Regional plot of the *SOST* region for type 2 diabetes (GCST006867)

gwas %>%
  filter(id == "GCST006867") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs1107747",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-t2d
#| fig-cap: QQ plot of the *SOST* region for type 2 diabetes (GCST006867)

gwas %>%
  filter(id == "GCST006867") %>%
  qq_plot()
```

:::

### HTN

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-htn
#| fig-cap: Regional plot of the *SOST* region for hypertension (UKB200021065)

gwas %>%
  filter(id == "UKB200021065") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-htn
#| fig-cap: QQ plot of the *SOST* region for hypertension (UKB200021065)

gwas %>%
  filter(id == "UKB200021065") %>%
  qq_plot()
```

:::

### AAC

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-aac
#| fig-cap: Regional plot of the *SOST* region for abdominal aortic calcification (PHS000930AAC)

gwas %>%
  filter(id == "PHS000930AAC") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-aac
#| fig-cap: QQ plot of the *SOST* region for abdominal aortic calcification (PHS000930AAC)

gwas %>%
  filter(id == "PHS000930AAC") %>%
  qq_plot()
```

:::

### LDL

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-ldl
#| fig-cap: Regional plot of the *SOST* region for LDL cholesterol (UKB30780)

gwas %>%
  filter(id == "UKB30780") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-ldl
#| fig-cap: QQ plot of the *SOST* region for LDL cholesterol (UKB30780)

gwas %>%
  filter(id == "UKB30780") %>%
  qq_plot()
```

:::

### HDL

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-hdl
#| fig-cap: Regional plot of the *SOST* region for HDL cholesterol (UKB30760)

gwas %>%
  filter(id == "UKB30760") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants",
    highlights_label = FALSE
  )
```

Note: rs72836561 is a missense variant in *CD300LG* that is known to be associated with lipid profiles [@surakka2015].  

#### QQ plot

```{r}
#| label: fig-qq-sost-hdl
#| fig-cap: QQ plot of the *SOST* region for HDL cholesterol (UKB30760)

gwas %>%
  filter(id == "UKB30760") %>%
  qq_plot()
```

Note: rs72836561 is a missense variant in *CD300LG* that is known to be associated with lipid profiles [@surakka2015].  

:::

### TG

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-tg
#| fig-cap: Regional plot of the *SOST* region for triglycerides (UKB30870)

gwas %>%
  filter(id == "UKB30870") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants",
    highlights_label = FALSE
  )
```

Note: rs72836561 is a missense variant in *CD300LG* that is known to be associated with lipid profiles [@surakka2015].  

#### QQ plot

```{r}
#| label: fig-qq-sost-tg
#| fig-cap: QQ plot of the *SOST* region for triglycerides (UKB30870)

gwas %>%
  filter(id == "UKB30870") %>%
  qq_plot()
```

Note: rs72836561 is a missense variant in *CD300LG* that is known to be associated with lipid profiles [@surakka2015].  

:::

### APOA

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-apoa
#| fig-cap: Regional plot of the *SOST* region for apolipoprotein A-I (UKB30630)

gwas %>%
  filter(id == "UKB30630") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants",
    highlights_label = FALSE
  )
```

Note: rs72836561 is a missense variant in *CD300LG* that is known to be associated with lipid profiles [@surakka2015].  

#### QQ plot

```{r}
#| label: fig-qq-sost-apoa
#| fig-cap: QQ plot of the *SOST* region for apolipoprotein A-I (UKB30630)

gwas %>%
  filter(id == "UKB30630") %>%
  qq_plot()
```

Note: rs72836561 is a missense variant in *CD300LG* that is known to be associated with lipid profiles [@surakka2015].  

:::

### APOB

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-apob
#| fig-cap: Regional plot of the *SOST* region for apolipoprotein B (UKB30640)

gwas %>%
  filter(id == "UKB30640") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-apob
#| fig-cap: QQ plot of the *SOST* region for apolipoprotein B (UKB30640)

gwas %>%
  filter(id == "UKB30640") %>%
  qq_plot()
```

:::

### FRAC

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-frac
#| fig-cap: Regional plot of the *SOST* region for fracture (GCST006980)

gwas %>%
  filter(id == "GCST006980") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-frac
#| fig-cap: QQ plot of the *SOST* region for fracture (GCST006980)

gwas %>%
  filter(id == "GCST006980") %>%
  qq_plot()
```

:::

### BMD

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-sost-bmd
#| fig-cap: Regional plot of the *SOST* region for heel bone mineral density (GCST006979)

gwas %>%
  filter(id == "GCST006979") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-sost-bmd
#| fig-cap: QQ plot of the *SOST* region for heel bone mineral density (GCST006979)

gwas %>%
  filter(id == "GCST006979") %>%
  qq_plot()
```

:::

:::

## Additional datasets

Genetic associations in the *SOST* region for the additional recent GWAS datasets in [@tbl-gwas-study-info].  

::: {.panel-tabset}

### CAD

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-add-sost-cad
#| fig-cap: Regional plot of the *SOST* region for coronary artery disease (GCST90132315)

gwas %>%
  filter(id == "GCST90132315") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

Note: the allele of rs851058 associated with lower coronary artery disease risk is also associated with lower mRNA expression of the *SOST* in [GTEx](https://gtexportal.org/home/snp/rs851058) [@gtex2020] and higher bone mineral density [@morris2019].  

#### QQ plot

```{r}
#| label: fig-qq-add-sost-cad
#| fig-cap: QQ plot of the *SOST* region for coronary artery disease (GCST90132315)

gwas %>%
  filter(id == "GCST90132315") %>%
  qq_plot()
```

Note: the allele of rs851058 associated with lower coronary artery disease risk is also associated with lower mRNA expression of the *SOST* in [GTEx](https://gtexportal.org/home/snp/rs851058) [@gtex2020] and higher bone mineral density [@morris2019].  

:::

### IS

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-add-sost-is
#| fig-cap: Regional plot of the *SOST* region for ischemic stroke (GCST90104535)

gwas %>%
  filter(id == "GCST90104535") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-add-sost-is
#| fig-cap: QQ plot of the *SOST* region for ischemic stroke (GCST90104535)

gwas %>%
  filter(id == "GCST90104535") %>%
  qq_plot()
```

:::

### CES

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-add-sost-ces
#| fig-cap: Regional plot of the *SOST* region for cardioembolic stroke (GCST90104536)

gwas %>%
  filter(id == "GCST90104536") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-add-sost-ces
#| fig-cap: QQ plot of the *SOST* region for cardioembolic stroke (GCST90104536)

gwas %>%
  filter(id == "GCST90104536") %>%
  qq_plot()
```

:::

### LAS

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-add-sost-las
#| fig-cap: Regional plot of the *SOST* region for large artery stroke	(GCST90104538)

gwas %>%
  filter(id == "GCST90104538") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-add-sost-las
#| fig-cap: QQ plot of the *SOST* region for large artery stroke	(GCST90104538)

gwas %>%
  filter(id == "GCST90104538") %>%
  qq_plot()
```

:::

### SVS

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-add-sost-svs
#| fig-cap: Regional plot of the *SOST* region for small vessel stroke	(GCST90104537)

gwas %>%
  filter(id == "GCST90104537") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-add-sost-svs
#| fig-cap: QQ plot of the *SOST* region for small vessel stroke (GCST90104537)

gwas %>%
  filter(id == "GCST90104537") %>%
  qq_plot()
```

:::

### T2D

::: {.panel-tabset .nav-pills}

#### Regional plot
```{r}
#| label: fig-reg-add-sost-t2d
#| fig-cap: Regional plot of the *SOST* region for type 2 diabetes (GCST90132183)

gwas %>%
  filter(id == "GCST90132183") %>%
  regional_plot(
    data = .,
    corr = ld_mat,
    top_marker = "rs66838809",
    highlights = pqtls$rsid,
    highlights_title = "Sclerostin associated variants"
  )
```

#### QQ plot

```{r}
#| label: fig-qq-add-sost-t2d
#| fig-cap: QQ plot of the *SOST* region for type 2 diabetes (GCST90132183)

gwas %>%
  filter(id == "GCST90132183") %>%
  qq_plot()
```

:::

:::

[^1]: The red and green dashed lines in the regional plots are the *p*-value thresholds $1 \times 10^{-6}$ (regional-wide significance threshold used in @zheng2023) and $5 \times 10^{-8}$ (genome-wide significance threshold), respectively.  
[^2]: The ischemic and cardioembolic stroke GWAS results from METASTROKE [@malik2016] used by @zheng2023 were replaced with those from MEGASTROKE [@malik2018] and the UK Biobank hypertension GWAS results from [OpenGWAS](https://gwas.mrcieu.ac.uk/datasets/ukb-b-14057/) used by @zheng2023 were replaced with those from [Pan-UKBB](https://pan.ukbb.broadinstitute.org/downloads) due to licensing restrictions. The GWAS of coronary artery calcification was not available either publicly or via application at the time of this analysis [@kavousi2022].  
