# Introduction

This code book contains the code used in our comment on "Lowering of Circulating Sclerostin May Increase Risk of Atherosclerosis and its Risk Factors: Evidence From a Genome-Wide Association Meta-Analysis Followed by Mendelian Randomization" by @zheng2023. These analyses use genetic variants in *cis* of the *SOST* gene (the gene that encodes sclerostin) that are associated with circulating sclerostin protein levels, known as protein quantitative trait loci (pQTL). We visualize the *SOST* gene region for each phenotype and perform Mendelian randomization (MR) and colocalization analyses for the cardiovascular events and risk factors tested using results from the same (or equivalent) genome-wide association studies (GWAS) as used by @zheng2023. Sensitivity Mendelian randomization analyses using European only samples where possible ([@sec-mr-eur]) and *cis* sclerostin pQTLs from UK Biobank [@sun2023] ([@sec-mr-ukbb]) were also performed.  

All of the code was written in `r fontawesome::fa(name = "r-project", fill = "#2468BB", height = "1em")` (version 4.3.1). This code book was written using Quarto (version 1.3.433).  

## Installation

The `data.table` and `tidyverse` packages were used to reading and processing data. The `Rsamtools` and `genetics` packages were used to query and process the 1000 Genomes data [@genomes2015]. The `geni.plots`, `heatmaply` and `reactable` packages are used to plot and present the results.  

```{r}
#| label: install
#| eval: false
install.packages("data.table", dependencies = TRUE)
install.packages("tidyverse")
install.packages("genetics")
install.packages("ieugwasr")
install.packages("BiocManager")
BiocManager::install("Rsamtools")
install.packages("remotes")
remotes::install_github("jrs95/geni.plots")
install.packages("heatmaply")
install.packages("reactable")
install.packages("downloadthis")
install.packages("tippy")
```

## Libraries

The libraries used throughout the code book are `dplyr`, `data.table`, `geni.plots` and `reactable`. Functions from other libraries are imported as and when needed. Additional functions used in the analyses are stored in `./scripts/00_functions.R` and are called using `source`.  

```{r}
#| label: setup
#| eval: false
library(data.table)
library(dplyr)
library(geni.plots)
library(reactable)
```

## Citation

Staley JR *et al*. Genetic data do not provide evidence that lower sclerostin is associated with increased risk of atherosclerosis: comment on the article by Zheng *et al*. Arthritis Rheumatol [2025;77(9):1280-1282](https://acrjournals.onlinelibrary.wiley.com/doi/10.1002/art.42751).  

## Licenses

The code is available under the GPL-3 license and the analysis results are available under the CC-BY-4.0 license.  

[![](https://img.shields.io/badge/Code-GPL--3-BD0000.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html) [![](https://img.shields.io/badge/Data-CC--BY--4.0-00AFD4.svg)](https://creativecommons.org/licenses/by/4.0/)  
